---
title: "Emacs-in"
author: 'Garbage Collector GNK'
date: '2022-02-04'
slug: []
categories: ["R", "Emacs"]
tags: ["R", "Emacs"]
---

# Using a free implementation of `vi` is but a penance: or I should use <kbd>SPC</kbd> more
This is the first attempt I have made to create a post from Emacs and also to
understand how blogdown actually works. The other first two posts I created, one
which was a garbage test post, and the other my first actual post, were made in
RStudio. Sinful!

1. Open Emacs (`emacsclient`)
2. <kbd>Meta</kbd><kbd>x</kbd> `R`
3. Load `blogdown` and `tidyverse` (even though I'm not doing any actual R work, so I didn't need to load the tidyverse. It's just habit)
4. ```{r}new_post(title="Emacs-in")```

I observe that without further arguments, the default format is markdown. I rename the file to make it an RMarkdown file:

<kbd>SPC</kbd> <kbd>f</kbd> <kbd>R</kbd>
<kbd>C-b</kbd> <kbd>C-b</kbd>
`R`
<kbd>Enter</kbd>

# Writing assistive functions in Emacs LISP
I would like to create a modified form of `evil-surround` so that I can easily
surround some keyboard key text with the HTML `kbd` tag. I can bind this to a
key, in some map (probably only `ess-r-mode-map`), or I might look into function
advice and create some template text.

For instance, rather than binding it to a key of its own, I might extend
`evil-surround` such that when a work is surrounded by KBD, like `kbdM-xkbd` the
surrounding tag is converted to the actual HTML: `<kbd>` and `</kbd>`.

Both would be fine.

I'd prefer to use
[KeyMason](http://www.miragecraft.com/projects/keymason.html)'s wonderful CSS
and JavaScript to supersede the CSS I got from Mozilla Developer Network (MDN is
indispensible when you're learning anything related to the web; the writing is
excellent). KeyMason shows individual keys very nicely.

When including this typographic flair, how should key combinations be displayed?
I think displaying it like a word, where each key is not separated by a space,
is preferrable to using some additive syntax:

<p><kbd>Shift</kbd>$+$<kbd>I</kbd></p>
<p><kbd>a</kbd> <kbd>m</kbd></p>
<p><kbd>n</kbd> <kbd>o</kbd> <kbd>t</kbd></p>
<p><kbd>a</kbd></p>
<p><kbd>n</kbd> <kbd>u</kbd> <kbd>m</kbd> <kbd>b</kbd> <kbd>e</kbd> <kbd>r</kbd></p>

The above is an anti-example. There are also already conventions for displaying
things that should be entered by keyboard: `like this`. Longer key sequences,
regular prose, should *never* be represented as keyboard keys. What should be
represented as key sequences are command sequences, such as

<p><kbd>g</kbd> <kbd>w</kbd></p>

to summon `evil-fill` (which wraps text to the `fill-column` variable width of
characters per line).

> `fill-column`

> /Documentation:/ Column beyond which automatic line-wrapping should happen. It
> is used by filling commands, such as `fill-region' and `fill-paragraph', and
> by `auto-fill-mode', which see. See also `current-fill-column'. Interactively,
> you can set the buffer local value using C-x f.

At least for my own purposes, creating a style guide for writing about key
sequences and adopting a more explicit visual style for keyboard keys will help
me discuss things like what button I have defined in KMonad or some sequence of
keys that I press to run a command to get my work done when I am writing
educational content.

<!-- # Progress towards the stated goals -->
<!-- ## Emacs LISP function -->
<!-- ## Writing style -->
<!-- ## Visual style -->

## How I use evil to just get by quicker
Without taking the time to write some Emacs LISP, debug it, bind it, and hope that I actually use it often enough to justify the time, here is how I get by:

- Put each key label on its own line
- Highlight the lines using visual mode <kbd>Shift</kbd><kbd>v</kbd>
- <kbd>Shift</kbd><kbd>I</kbd> then type the opening `<kbd>` tag
- Highlight the lines again
- <kbd>Shift</kbd><kbd>a</kbd> then type the closing `</kbd>` tag

Done! Modal editing power brought to Emacs. Evil is wonderful, but it's so sinful, ain't it?

# How blogdown (sort of) works
When `blogdown::new_post()` is run, at least in Emacs running as a daemon, the
iESS buffer will be occupied by the function's execution until you are done
editing that post. When you *are* done editing, you should use <kbd>C-x</kbd>
<kbd>#</kbd> to finish editing the post with Emacs `server-edit`. You might miss
that you are supposed to do this if you are not looking at the echo area or the
messages buffer when you run create a new post with this function.

```{r eval=FALSE}
> new_post(title="Emacs-in")
Waiting for Emacs...
[1] "/home/garbagecollectorgnk/code/personalBlogdown/content/post/2022-02-04-emacs-in/index.md"
```

> `describe-function` (`C-h f` or `SPC H d f`): `server-edit`

> Switch to next server editing buffer; say "Done" for current buffer. If a
> server buffer is current, it is marked "done" and optionally saved. The buffer
> is also killed if it did not exist before the clients asked for it. When all
> of a client’s buffers are marked as "done", the client is notified.

<kbd>C-x</kbd> <kbd>#</kbd>

This is necessary when you're using Emacs as a daemon because
`blogdown::new_post()` calls a private function, `blogdown:::open_file(file,
open)` where `file` is a YAML-good filename made from the title of the post if a
file is not specified. `open_file` falls back to `base::edit.file()` using
`tryCatch` if you're not using RStudio (the call to `rstudioapi::navigateToFile`
will fail), and `edit.file` simply uses the `$EDITOR` environment variable,
which in my case will be emacsclient.

The next thing `blogdown::new_post()` appears to do is some background work to
create the directory for the new post in `./content/post/`; I didn't read
`blogdown:::new_content()` very closely, after reading the terse new post
function I didn't feel like it, but that's probably where the calls to this
occur becaus they aren't in new post.

## Uploading
To upload the new post to our Hugo website using Netlify (if you're using these) we *may* need to knit our document. We may need to do so at least if we're using RMarkdown, aren't pushing that to the repository, or haven't changed the default extension from `.md` to `.Rmd` like [Yihui hints](https://yihui.org/en/2021/05/blogdown-knit-save/#4-consistent-behavior-with-md-posts) is common to do but unnecessary if you're not even evaluating any R code (if you're always using `eval=FALSE}`).

Perhaps I made the wrong decision to blindly choose Rmd as well…

Anyways, using `blogdown::build_site()`, `check_site()`, and `rmarkdown::knitr()` we get our HTML and then run `check_site()` again. No errors, no warnings, no TODOs. Commit (no sin by) using Magit, push to remote.origin with SSH and we should see the post become live on the Netlify site in a just a moment!

I do recommend setting up SSH keys. They make life a lot easier, and more secure. When I made the first two posts (a test and "keyboardingIntoTheFuture") in RStudio I was annoyed that even though I used the recommend functions `use_github` (I think) it wouldn't rememeber my personal token for more than about half an hour and I would need to enter it again upon a push. With an SSH key this will never occur again. :) At least if you get your remote URL set correctly: it needs to be `ssh://git@github.com/USERNAME/REPOSITORY`, you cannot just change from `https` TO `ssh` if you have a remote configured already. Be sure to type `ssh://git@` when you're prending and ensure the URL is then of the form just mentioned.

…this is why we always run `check_site()`!:

```{r eval=FALSE}
● [TODO] Found 1 R Markdown file to update by re-rendering:

  content/post/2022-02-04-emacs-in/index.Rmd

  To update a file, re-knit or use blogdown::build_site(build_rmd = 'timestamp')
```

```{r eval = FALSE}
build_site(build_rmd = 'timestamp')
```

We do *not* need to knit ourselves, we can just build (the) site using blogdown and forget about it, specially so if there is no R code to evaluate (a simple Rmd to md conversion alone using knitr without any real knitting).
